#Practical Machine Learning Course Project
####Author: Chong Kean Meng  

##Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The project goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

```{r, echo=FALSE}
suppressWarnings(library(caret))
suppressWarnings(library(rpart))
suppressWarnings(library(rpart.plot))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(rattle))
suppressWarnings(library(randomForest))
suppressWarnings(library(gbm))
suppressWarnings(library(plyr))
```

##Data Processing
```{r}
#Acquire Training and Test Data
if (!file.exists("pml-training.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                  destfile = "pml-training.csv")
}
if (!file.exists("pml-testing.csv")) {
    download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")
}

#Load data
train <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"))                                                
test <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"))
dim(train)
dim(test)
```
Clean the data by removing missing values and unused variables.
```{r}
sum(complete.cases(train))

##Remove columns that contain NA missing values.
train <- train[, colSums(is.na(train)) == 0] 
test <- test[, colSums(is.na(test)) == 0] 

## Remove columns that do not contribute to the accelerometer measurements.
classe <- train$classe
trainRemove <- grepl("^X|timestamp|window", names(train))
train <- train[, !trainRemove]
trainCleaned <- train[, sapply(train, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(test))
test <- test[, !testRemove]
testCleaned <- test[, sapply(test, is.numeric)]
dim(trainCleaned)
dim(testCleaned)

#Verify there is no zero covariates as they are not likely to be good predictors
myDataNZV <- nearZeroVar(train, saveMetrics=TRUE)
myDataNZV
```

##Cross Validation
The pml.training data will be used to train the model.  To estimate the pml.testing data accuracy, the pml.training data is splitted into 60% (Training set) and 40% (Test set).
```{r}
set.seed(12345) #for reproducible purpose
inTrain <- createDataPartition(y=trainCleaned$classe, p=0.6, list=FALSE)
mytrain <- trainCleaned[inTrain, ]
mytest <- trainCleaned[-inTrain, ]
```
```{r}
dim(mytrain)
```
```{r}
dim(mytest)
```

##Determine Prediction Model
####Comparison of prediction models - Regression and Classification Trees and Random Forest
###Regression and Classification Trees 
```{r}
modelRpart <- rpart(classe ~ ., data=mytrain, method="class")
prp(modelRpart)
predictRpart <- predict(modelRpart, mytest, type = "class")
confusionMatrix(mytest$classe, predictRpart)
```
Decision Tree has an accuracy rate of **0.7267** and Out of Sample Error rate is **0.2733** 

###Random Forest
```{r}
controlRf <- trainControl(method="cv", 10)
modelRf <- train(classe ~ ., data=mytrain, method="rf", trControl=controlRf, ntree=250)
modelRf
predictRf <- predict(modelRf, mytest)
confusionMatrix(mytest$classe, predictRf)
```
Random Forest has an accuracy rate of **0.9927** and Out of Sample Error rate is **0.0073** 

> Between the two models, Random Forest is a better Prediction Model.  

##Predicting Test Data Set using Random Forest
```{r}
outcome = predict(modelRf, testCleaned)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(outcome)
```
